---
title: "Canadian Weather Data"
author: "fda.clust Development Team"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Canadian Weather Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(fda)
library(fda.usc)
library(fda.clust)
```

## **Data Preparation**

We will first extract the daily temperature and precipitation data and organize it as **fdata** objects. We will also prepare a data frame containing metadata for each station, such as its name, region, coordinates, and basic descriptive statistics (mean, min, and max) for temperature and precipitation.

```{r data-preparation, fig.width=7, fig.height=5, fig.cap = "Preparation of temperature and precipitation data for clustering."}
# Load the Canadian Weather data
data("CanadianWeather", package = "fda")

# Extract daily temperature and precipitation data and transpose them
temp_data <- t(CanadianWeather$dailyAv[, , "Temperature.C"])
prec_data <- t(CanadianWeather$dailyAv[, , "Precipitation.mm"])

# Create fdata objects for temperature and precipitation
n <- nrow(temp_data)
p <- ncol(temp_data)
argvals <- seq(1, 365, length.out = p)
temp <- fdata(temp_data, argvals = argvals)
prec <- fdata(prec_data, argvals = argvals)

# Calculate descriptive statistics for temperature and precipitation
station_names <- names(CanadianWeather$region)
coordinates <- CanadianWeather$coordinates
corrected_longitude <- ifelse(coordinates[, 2] > 180, coordinates[, 2] - 360, coordinates[, 2])

mean_temp <- rowMeans(temp$data, na.rm = TRUE)
min_temp <- apply(temp$data, 1, min, na.rm = TRUE)
max_temp <- apply(temp$data, 1, max, na.rm = TRUE)
mean_prec <- rowMeans(prec$data, na.rm = TRUE)
min_prec <- apply(prec$data, 1, min, na.rm = TRUE)
max_prec <- apply(prec$data, 1, max, na.rm = TRUE)

# Create a data frame with descriptive statistics
df <- data.frame(
  Station = station_names,
  Region = CanadianWeather$region,
  MeanTemp = mean_temp,
  MinTemp = min_temp,
  MaxTemp = max_temp,
  MeanPrec = mean_prec,
  MinPrec = min_prec,
  MaxPrec = max_prec,
  Latitude = coordinates[, 1],
  Longitude = corrected_longitude
)

# Create ldata object
weather <- ldata(
  df = df,
  temp = temp,
  prec = prec
)
```


## **Clustering Functional Data**

We will now cluster the temperature data using algorithms from **fda.clust**, including **fKmeans**, **fDBSCAN**, and **fHclust**.

### **fKmeans**

We will cluster the temperature data using **fKmeans**  with 3 clusters.from **fda.clust**. 

```{r fkmeans-clustering, fig.width=7, fig.height=5, fig.cap = "Clustering of temperature curves using fKmeans with 3 clusters."}
# Apply fKmeans to cluster temperature curves
set.seed(123) 
kmeans_result <- fkmeans(temp, clusters = 3)
plot(temp, col = kmeans_result$cluster, main = "fKmeans Clustering of Temperature Data")
```

### **fDBSCAN**

We use the optimization function from **fda.clust** to select appropriate values for `eps` and `minPts`.

```{r fdbscan-clustering, fig.width=7, fig.height=5, fig.cap = "Clustering of temperature curves using fDBSCAN."}
# Optimize parameters for fDBSCAN
set.seed(123)
optimal_params <- optim.fdbscan(temp)$optimal

# Apply fDBSCAN with optimal parameters
dbscan_result <- fdbscan(temp,
                  eps = optimal_params$eps,
                  minPts = optimal_params$minPts)

plot(temp, col = dbscan_result$cluster, main = "fDBSCAN Clustering of Temperature Data")
```

### **fHclust**

We use hierarchical clustering with the **Ward's method**.

```{r fhclust-clustering, fig.cap = "Hierarchical clustering dendrogram of weather data."}
# Apply hierarchical clustering to temperature data
dist_matrix <- metric.lp(temp$data)
hclust_result <- fhclust(dist_matrix, method = "ward.D2")
plot(hclust_result, labels = df$Station, main = "Hierarchical Clustering Dendrogram")
```

## **Visualizing Weather Stations on the Map**

```{r plot-map, eval=FALSE, fig.cap = "Map of weather stations in Canada with their locations highlighted."}
library(maps)
plot(NA, xlim = c(-150, -50), ylim = c(40, 90), type = "n", xlab = "Longitude", ylab = "Latitude", main = "Weather Stations in Canada")
map("world", regions = "Canada", fill = TRUE, col = "lightblue", add = TRUE)
points(df$Longitude, df$Latitude, pch = 21, bg = "red", col = "black", cex = 1.2)
text(df$Longitude, df$Latitude, labels = df$Station, pos = 4, cex = 0.7)
```

